---
title: "Lab2_Wei"
output: html_document
date: "2024-09-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message= FALSE,
                      warning= FALSE,
                      rm(list=ls()),
                      options=(scipen=999)
                      )
```

```{r}
library(nycflights13)
library(tidyverse)
```

```{r}
flights
```


### 5.2.4
1-1. Find all flights that had an arrival delay of two or more hours. What proportion of all flights was this
group? How does this number compare with flights that were delayed between one and two hours (inclusive)?

- The proportion of flights that had an arrival delay of two or more hours is 0.03028719. The proportion of flights that were delayed between one and two hours is 0.05428831. The proportion of flights delayed between one and two hours is higher than those delayed by two or more hours, which suggests that shorter delay might be more common.  

```{r}
# Filter flights that had an arrival delay of two or more hours.
delay2<-flights %>%
  filter(arr_delay>=120)  

delay2
```


```{r}

# Calculate the proportion of flights by two or more hours
prop_delay<-nrow(delay2)/nrow(flights)

# Filter flights that had an arrival delay of between one and two hours
delay1<-flights %>%
  filter(60<=arr_delay & arr_delay<=120)

# Calculate the proportion of flights between 1 and 2 hours
prop_delay2<-nrow(delay1)/nrow(flights)

prop_delay
prop_delay2

```


1-2. Find all flights that flew to Houston (IAH or HOU). Was this more or fewer than the number of flights
to Atlanta (ATL)? What proportion of all flights went to Houston? Atlanta?

- The number of flights to Atlanta which is 17,215 is greater than flights to Houston which is 9313. Also Atlanta has higher shares of flights compared to Houston as a destination. 

```{r}
# All flights that flew to Houston (IAH or HOU)
flights_houston<-flights %>%
  filter(dest %in% c("IAH", "HOU"))

# Compute the number of flights to Houston (IAH or HOU) and its proportion 
num_flights_houston=nrow(flights_houston)
prop_houston=num_flights_houston/nrow(flights)

# All flights that flew to Atlanta (ATL)
flights_atl<-flights %>%
  filter(dest == "ATL")

# Compute the number of flights to Atlanta (ATL) and its proportion 
num_flights_atl=nrow(flights_atl)
prop_ALT=num_flights_atl/nrow(flights)

# Display
num_flights_atl
num_flights_houston
prop_houston
prop_ALT
```


1-3. What proportion of flights to Atlanta were operated by Delta Airlines? What might explain the size of
this proportion? Compare it with the proportion of United flights arriving at Chicago O’Hare.

-Delta operates over 61% of all flights to Atlanta, which can be explained by Atlanta being Delta's largest hub and its headquarters. In comparison, United handles 40% of flights arriving at Chicago O’Hare.


```{r}
# Flights to Atlanta operated by Delta Airlines
flights_delta_atl <- flights %>%
  filter(dest == "ATL"& carrier == "DL")

# Compute proportion of flights to Atlanta operated by Delta Airlines
prop_delta_atlanta <- nrow(flights_delta_atl) / num_flights_atl

# Flights to Chicago operated by United Airlines
flights_united_ord <- flights %>%
  filter(dest == "ORD" & carrier == "UA")

# Compute proportion of flights to Atlanta operated by Delta Airlines
prop_united_ord <- nrow(flights_united_ord) / nrow(flights %>%
  filter(dest == "ORD"))

prop_delta_atlanta
prop_united_ord
```


1-7. Find all flights that departed between midnight and 6am (inclusive). Compare this number with the number departing between noon and 6 pm (inclusive).

- There is a significant difference in the number of flights during these two time periods. Only 9,344 flights departed between midnight and 6am, while 121,141 flights departed between noon and 6pm. This shows that the majority of flights occur during the afternoon hours, likely due to higher passenger demand and more favorable operational conditions during the day.
```{r}
# Flights departing between midnight and 6 am
flights_midnight <- flights %>%
  filter(0<=dep_time & dep_time <= 600)
num_midnight=nrow(flights_midnight)
# Flights departing between noon and 6 pm
flights_noon <- flights %>%
  filter(1200<=dep_time & dep_time <= 1800)
num_noon=nrow(flights_noon)

num_midnight
num_noon
```

### 5.3.1
2. Sort flights to find the most delayed flights. Find the flights that left earliest. In your written answer provide the details of the single most delayed flight and the flight that left most ahead of schedule, including flight number, destination, and delay (positive or negative) in hours and minutes.

- The most delayed flight was flight 51 to Honolulu (HNL), which was delayed by 21 hours and 41 minutes.
- The flight that left most ahead of schedule was flight 97 to Denver (DEN), departing 43 minutes early.
```{r}
# Find the most delayed flights
most_delay<-flights %>%
  arrange(desc(dep_delay)) %>%
  select(dep_delay,dest,flight)  %>%
  slice_head()

# Find the flights that left earliest
earliest_flight<-flights %>%
  arrange(dep_delay) %>%
  select(dep_delay,dest,flight)  %>%
  slice_head()

# Convert delay from minutes to hours and minutes
most_delay <- most_delay %>%
  mutate(
    delay_hours = floor(dep_delay / 60),
    delay_minutes = dep_delay %% 60
  )

earliest_flight <- earliest_flight %>%
  mutate(
    delay_hours = floor(dep_delay / 60),
    delay_minutes = dep_delay %% 60
  ) 

most_delay
earliest_flight
  
```




### 5.4.1
2. What happens if you include the name of a variable multiple times in a select() call?

-Even though the name of a variable is included multiple times in select(), the variable only appears once. 
```{r}
flights %>%
  select(dep_delay, month, dep_delay)
```

3. What does the one_of() function do? Why might it be helpful in conjunction with this vector?
vars <- c("year", "month", "day", "dep_delay", "arr_delay")

- one_of() function select columns from a data frame based on a character vector of column names. This method is dynamic and allows for flexible selection based on an external list of column names.


```{r}
vars <- c("year", "month", "day", "dep_delay", "arr_delay")

flights %>%
  select(one_of(vars))
```

